Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-09 ba_resubmit/kundinger_resubmit.tex
cleaning up

\textcolor{red}{Consider two databases $\bf{X}_1$ and $\bf{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi  & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + j \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$

@@

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in section~\ref{bayes-estimate}.

@@

\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi  & z_j \leq n_1; \\
		1-\pi &  z_j  = n_1 + j \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.

@@

$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z}))$$
$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z}))$$
$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{z_j, j}  & z_j \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & z_j   = n_1 + j. \\
\end{cases},$$
where 
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
Derivations for these full conditionals can be found in Section ~\ref{app:derivations}.

@@

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
\end{cases}\] 
Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.

@@

We seek to minimize the posterior expected loss, given by 
$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$

Here, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule

%$\theta_R = \infty, \theta_{10} = 1, \theta_{01} = 1, \theta_{11} = 2$ ,

\[\hat{Z}_j =\begin{cases} 
	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
	0,  & \text{otherwise}. \\
\end{cases}\]

For a more in-depth explanation of this function and the induced Bayes
estimate, see \cite{sadinle_bayesian_2017}.

@@

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}.$$
For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics. Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.

@@

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_2} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes

@@

Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters). Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have

@@

Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\). Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have

%As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with

\begin{align*}
	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&=
	\begin{cases} 
		w_{z_j, j}  & z_j \leq n_1; \\
		1 &  z_j  = n_1 + j \\
	\end{cases},\\
\end{align*}
where
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$

With the likelihood in this form, we derive an expression for the posterior distribution of $Z_j$:
\begin{align*}
	p\left(Z_j^{(s+1)}|\Gamma_{.j}, \Phi, \bm{Z^{(s)}}, \pi\right) &\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \Phi) P(Z_j^{(s+1)}|\pi) \\
	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j). \\
	&= \begin{cases} 
		\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
		1-\pi &  z_j  = n_1 + j. \\
	\end{cases}
\end{align*}

For more direct comparability with the method from \cite{sadinle2018bayesian}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.

>>>>

\textcolor{red}{Consider two databases $\bf{X}_1$ and $\bf{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts.} In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$

@@

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section~\ref{bayes-estimate}.

@@

\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
		1-\pi, &  z_j  = n_1 + j; \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.

@@

$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{z_j, j},  & z_j \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
\end{cases},$$
where 
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.

@@

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0,  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
\end{cases}\] 
Here, \(\theta_R\) is the loss from not making a decision on the linkage status, \(\theta_{10}\) is the loss from a false non-match, \(\theta_{01}\) is the loss from a false match, and \(\theta_{11}\) is the loss from the special case of a false match in which the record has a true match other than the one estimated by the model.

@@

We seek to minimize the posterior expected loss, given by 
$$\mathbb{E}[L(\hat{\bm{Z}}, \bm{Z})|\Gamma^{obs}] = \sum_{\bm{Z}} L(\hat{Z_j}, Z_j)P(\bm{Z}|\Gamma^{obs}).$$
In general, we set $(\theta_{10}, \theta_{01}, \theta_{11}, \theta_R) = (1, 1, 2, \infty)$ inducing the intuitive decision rule
$$\hat{Z}_j =\begin{cases} 
	i,  & \text{if } P(Z_j = i |\Gamma) > \frac{1}{2}; \\
	0,  & \text{otherwise}. \\
\end{cases}$$
For a more in-depth explanation of this function and the induced Bayes
estimate, see \cite{sadinle_bayesian_2017}.

@@

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2},$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||,$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}.$$
For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{p}\}\right\}.$ As we will show in Section \ref{efficient-posterior}, all posterior calculations are conducted with these sufficient statistics. Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for posterior inference.

@@

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes

@@

In order to define the full conditional for $\bm{Z}$, we must provide the posterior distribution for $\pi$, and provide a more tractable expression for the portion of the likelihooh relevant to a particular record $j \in \bm{X}_2$. Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bm{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + j)$ encoded by $\bm{Z}$ (and hyperparameters). Thus we use $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have

@@

Let \(\Gamma_{.j}\) denote the set of \(n_1\) comparison vectors with \(j \in \bm{X_2}\). We have
%As a function of \(Z_j\), \(p(\Gamma_{.j}|Z_j, \Phi) = \mathcal{L}(Z_j|\Gamma_{.j}, \Phi)\) is a discrete distribution with
\begin{align*}
	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
	&\propto \prod_{i=1}^{n_1}\frac{\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}{\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}}\\
	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&=
	\begin{cases} 
		w_{z_j, j}  & z_j \leq n_1; \\
		1 &  z_j  = n_1 + j \\
	\end{cases},\\
\end{align*}
where
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$

With the likelihood in this form, we derive an expression for the posterior distribution of $\bm{Z}$. Because we sample each \(Z_j\) independently of all other \(Z_{j'}\), we use only the full conditional for an individual \(Z_j\):
\begin{align*}
	p\left(Z_j^{(s+1)}|\Gamma_{.j}, \Phi, \bm{Z^{(s)}}, \pi\right) &\propto p(\Gamma_{.j}| Z_j^{(s+1)}, \Phi) P(Z_j^{(s+1)}|\pi) \\
	&\propto \left(\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + I(z_j = n_1 + j) \right) \left (\pi\sum_{i=1}^{n_1}\frac{1}{n_1}I(z_j = i) + (1-\pi)I(z_j = n_1 + j) \right) \\	
	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{z_j, j}I(z_j = i) + (1-\pi)I(z_j = n_1 + j). \\
	&= \begin{cases} 
		\frac{\pi}{n_1}w_{z_{j}, j}   & z_j \leq n_1; \\
		1-\pi &  z_j  = n_1 + j. \\
	\end{cases}
\end{align*}

For more direct comparability with the method from \cite{sadinle_bayesian_2017}, we avoid sampling $\pi$ directly, and instead integrate over it in the final full conditional for $Z_j$.
